\chapter*{Aknowledgements}

The content here presented was produced during the last academic year of my bachelor's degree in Mathematics at the Institute of Mathematics and Statistics of the University of São Paulo (IME-USP). This material represents a very important end of cycle for me, and therefore, I need to thank those responsible for not letting me give up on everything.

First of all, I couldn't forget to mention my dear parents: \textsc{Ricardo Mezzovilla Gonçalves} and \textsc{Manoelina Pauzner}. They are the ones who know how passionate I am about mathematics, and in particular, my Mother has always shown the greatest interest, asking me to explain different and diverse topics. The opportunity to study was granted to me by them, in addition to always encouraging me.

During my studies, I also worked in parallel on completely orthogonal topics, to which I've been struggling to balance those bad rocks. Giving up was constantly on my mind. To the love of my life, \textsc{Maiara dos Anjos Justino}, I must thank for preventing me from giving up, giving me the strength to keep going, in a way only she would be able to lift me up. Only she saw my journeys against the tide, wearing me down not only physically but psychologically. Even if it's cliche, I don't know what I would do without her presence all these years.

I must also thank my advisor \textsc{Severino Toscano do Rego Melo} who agreed to embark on this journey with me, which in several moments I'had seen my self lost and without his help, I wouldn't be able to continue.

I would like to give honorable mentions to my friends, which, directly or indirectly were responsible for my journey up here: \textsc{Francisco Felipe Chagas Rangel}, \textsc{Zeus Santos Nazaro da Silva}, \textsc{Isabella Matias}, \textsc{Guilherme Cerqueira Gonçalves} and \textsc{Andrey Luan Gomes Contel}. Fortunately, I have a large list of friends I know I can count on, and I would like to name all my friends from College, Work and the other four corners of the world, but I have a terrible memory and I don't want to go into debt with anyone.